Implications of afforestation policies for global markets

### 2023-08-07 paper.Rmd
Update paper.Rmd

In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023]. These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emissions [@roe2019]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation. There is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019]. This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].  In many cases, these results suggest massive losses in pasture and cropland. This puts into question the feasibility of attaining these results, as forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland. Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017]. Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

>>>>

In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4% of the land [@fao2023]. These forests play an important role in maintaining and growing the land-based carbon sink by currently sequestering about 30% of GHG emissions [@roe2019]. This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation. As such, there is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020]. Many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, also rely to a significant degree on afforestation projects [@roe2019].

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. The population growth that is especially high, at present, in Africa also puts pressure on policymakers to allow greater conversion of forestland into farmland. Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
